---
output:
  html_document: default
  pdf_document: default
---

Title My Reproducible Research Project1 

```{r}
library(ggplot2)
library(data.table)
library(Hmisc)
```
What is mean total number of steps taken per day?

Preprocessing Data

1. load the data (ie. read.csv())
```{r}
Data <- read.csv("activity.csv")
str(Data)
```
2. Process and transform data if needed
```{r}
Data1 <- Data[complete.cases(Data),]
Tsteps <- aggregate(steps ~ date, Data1, sum)
names(Tsteps)[1] <- "Date"
names(Tsteps)[2] <- "Totalsteps"
```
3. Make a histogram of the total of steps taken per day
```{r}
hist(Tsteps$Totalsteps,
     col="red",
     main="Histogram of the total number of steps per day",
     xlab="The total number of steps per day",
     breaks =10)
```
4. Calculate and report the mean and median number of steps per day
```{r}
mean(Tsteps$Totalsteps)
median(Tsteps$Totalsteps)
```
What is the average daily activity pattern

1. Make a time series plot of 5-min interval and the average number of steps taken, averaged across all days
```{r}
Data2 <- Data[complete.cases(Data),]
MData1 <- aggregate(Data2$steps, by=list(Data2$interval), mean)
names(MData1)[1] <- "interval"
names(MData1)[2] <- "steps"
ggplot(MData1, aes(x=interval, y=steps)) + 
        labs(title="Sum of Steps By Interval", x="interval", y="steps")+
        geom_line(color="blue")
```
2. Which 5-min interval contains max numbers of steps?
```{r}
Interval <- MData1[which.max(MData1$steps),]
Interval
```
Imputing Missing Values

1. Calculate and report the total number of missing values in the dataset (i.e the total number of rows with NAs)
```{r}
Datamissing <- sum(is.na(Data))
Datamissing
```

2. Devise a strategy for filling in all of the missing values in the dataset.
My strategy: will use the mean interval steps for the 5-minute interval at a given interval.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
NewData <- Data
NewData <- NewData[complete.cases(NewData$steps),]
MeanByInterval <- aggregate(NewData$steps, by=list(NewData$interval), sum)
names(MeanByInterval)[1]="interval"
names(MeanByInterval)[2]="steps"
```
Impute Method- Attempt 2
```{r}
NewData1 <- Data
Datamissing <- is.na(NewData1$steps)
CleanDatamissing <- NewData1[!is.na(NewData1$steps),]
MeanVals <- tapply(CleanDatamissing$steps, CleanDatamissing$interval, mean, na.rm=TRUE, simplify=TRUE)
NewData1$steps[Datamissing] <- MeanVals[as.character(NewData1$interval[Datamissing])]
sum(Datamissing)
sum(is.na(NewData1$steps))
```

4. Make a histogram of the total number of steps taken per day and calculate and report the mean and median total number of steps taken per day.
```{r}
SumDataByDay <- aggregate(NewData1$steps, by=list(NewData1$date), sum)
names(SumDataByDay)[1]="date"
names(SumDataByDay)[2]="totalsteps"
head(SumDataByDay,20)
# Plot using ggplot
ggplot(SumDataByDay, aes(x=totalsteps))+
        geom_histogram(fill="maroon",binwidth=1000)+
        labs(title="Total Daily Steps", x="Steps", y="Frequency")
# Mean on New Data
mean(SumDataByDay$totalsteps)

#Median on New Data
median(SumDataByDay$totalsteps)
```
Yes, they are same with original.

Are there differences in activity patterns between weekdays and weekends?
```{r}
NewData1$weekday <- weekdays(as.Date.character(NewData1$date))
NewData1$weekend <- ifelse(NewData1$weekday=="Saturday"| NewData1$weekday=="Sunday","Weekend","Weekday")
head(NewData1,5)

MeanWEWD <- aggregate(NewData1$steps, by=list(NewData1$weekend, NewData1$interval), mean)
names(MeanWEWD)[1]="weekend"
names(MeanWEWD)[2]="interval"
names(MeanWEWD)[3]="steps"

ggplot(MeanWEWD, aes(x=interval, y=steps, color=weekend))+
        geom_line()+
        facet_grid(weekend ~.)+
        labs(title="Time Series Plot Of The 5-Minute Interval\nAveraged Across All Weekday Days or Weekend Days", x="interval", y="steps")
```
